---
title: "DisaggregateByFeedstock"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## load libraries
library(tidyr)
library(dplyr)
```

## Read csvs and tidy up the column names and data
```{r}
supplies <- read.csv("BiomassSuppliesbyState.csv", header= TRUE)
mapping <- read.csv("stateToRegionMapping.csv", header=TRUE)
use <- read.csv("biomassUseByZone.csv", header = TRUE)

## change colnames as needed
colnames(supplies)
colnames(mapping)
colnames(mapping) <- c("state", "egrid_agg", "allocator")
colnames(use)
colnames(use) <- c("Fuel", "Run.Name", "Year", "egrid_agg", "Dry.Tons.Use.egrid_agg")

## clean up state names (they don't match between the tables because of capitalization)
supplies <- supplies %>% 
  mutate(State = tolower(State))

### check that state names are the same (that the supply curve states are in the mapping table)
unique(supplies$State) %in% unique(mapping$state)

## remove commas and convert factors to numeric
supplies$Dry.Tons <- as.numeric(as.character(gsub(",", "", supplies$Dry.Tons)))
```

## Join mapping to supplies table
This gets the mapping[alloctator] column into the supplies table 
```{r}
supplies_zones <- left_join(supplies, mapping, by = c("State" = "state"))
nrow(supplies_zones)
nrow(supplies)

## multiply the Dry.Tons by the allocator
supplies_zones <- supplies_zones %>% 
  mutate(Dry.Tons.Alloc = Dry.Tons*allocator)
```

## Prep use table
```{r}
use <- use %>% 
  ## separate the resource form from the price
  separate(Fuel, c("Resource.Form", "Biomass.Price"), sep = "_", remove = TRUE) %>% 
  ## Isolate resource form by deleting "biomass primary - "
  mutate(Resource.Form = gsub("biomass primary - ", "", Resource.Form))

unique(use$Resource.Form)
nrow(use)

## delete Resource.Form == "corn" rows
use <- use %>% 
  filter(Resource.Form != "corn")

##check that rows were removed
unique(use$Resource.Form)
nrow(use)
```

## Prep supplies_zones table: 
1) remove unnecessary years
2) duplicate 2040 for 2045 and 2050
```{r}
unique(supplies_zones$Year)
unique(use$Year)

## in supplies_zones: remove all unnecessary years (in between the five year increments)
supplies_zones_sub <- supplies_zones %>% 
  filter(Year %in% unique(use$Year))

nrow(supplies_zones_sub)

## duplicate 2040 for 2045 and 2050 in supply curve
supplies_zones_sub_dup <- rbind(supplies_zones_sub, 
                            supplies_zones_sub %>% 
                              filter(Year == 2040) %>%
                              mutate(Year = 2045))

supplies_zones_sub_dup <- rbind(supplies_zones_sub_dup, 
                            supplies_zones_sub_dup %>% 
                              filter(Year == 2040) %>%
                              mutate(Year = 2050))

nrow(supplies_zones_sub_dup)
unique(supplies_zones_sub_dup$Year)

## Check that it has the right number of rows
nrow(filter(supplies_zones_sub_dup, Year == 2040))*2 + nrow(supplies_zones_sub) == nrow(supplies_zones_sub_dup)

#write.csv(supplies_zones_sub_dup, "supplies_zones_sub_dup.csv")
```
## Calculate allocation proportions in supplies_zones_sub_dup 
```{r}
sum.Dry.Tons.Alloc <- supplies_zones_sub_dup %>% 
  group_by(Biomass.Price, Resource.Form, Year, egrid_agg) %>%
  summarise(sum.Dry.Tons.Alloc = sum(Dry.Tons.Alloc))

## join sums to supplies_zones_sub_dup
supplies_zones_sub_dup <- left_join(supplies_zones_sub_dup, sum.Dry.Tons.Alloc, by = c("Biomass.Price" = "Biomass.Price", "Resource.Form" = "Resource.Form", "egrid_agg" = "egrid_agg", "Year" = "Year"))

## Divide Dry.Tons.Alloc by sum.Dry.Tons.Alloc
supplies_zones_sub_dup <- supplies_zones_sub_dup %>%
  mutate(use.allocator = Dry.Tons.Alloc/sum.Dry.Tons.Alloc)

#write.csv(supplies_zones_sub_dup, "supplies_zones_sub_dup_useAlloc.csv")
```

## join supplies_zones_sub_dup with use table
This allow the use.allocator to be applied to the Dry.Tons.Use.egrid_agg (selected tons of biomass in each scenario) to calculate the per state and resource allocation.
```{r}
# make Resource.Form in supplies table lowercase to match use table
supplies_zones_sub_dup <- supplies_zones_sub_dup %>%
  mutate(Resource.Form = tolower(Resource.Form))
unique(supplies_zones_sub_dup$Resource.Form)
unique(use$Resource.Form)

## ERR, it's "woody" in supplies table and "wood" in the use table! replace "wood" with "woody" in use table in the Resource.Form column
use <- use %>%
  mutate(Resource.Form = replace(Resource.Form, Resource.Form == "wood", "woody"))

unique(use$Resource.Form)

## ALSO, Biomass.Price in supplies table is a character. make it numeric:
supplies_zones_sub_dup$Biomass.Price <- as.numeric(supplies_zones_sub_dup$Biomass.Price)
use$Biomass.Price <- as.numeric(use$Biomass.Price)

## Now we can do the join:
supplies_zones_sub_dup_use <- left_join(supplies_zones_sub_dup, use, by = c("Biomass.Price" = "Biomass.Price", "Resource.Form" = "Resource.Form", "egrid_agg" = "egrid_agg", "Year" = "Year"))

## check that this join was done correctly by ensuring the number of additional rows (multiplied) equal the number of Run.Name scenarios
length(unique(use$Run.Name)) == nrow(supplies_zones_sub_dup_use)/nrow(supplies_zones_sub_dup)

```

## Calculate the disaggregated biomass usage numbers by lowest dimension (resource within each state), saved as column, Dry.Tons.Use.State
```{r}
## Convert all NaNs in the user.allocator (which are 0/0) to 0
is.nan.data.frame <- function(x)
do.call(cbind, lapply(x, is.nan))

supplies_zones_sub_dup_use[is.nan.data.frame(supplies_zones_sub_dup_use)] <- 0

supplies_zones_sub_dup_use <- supplies_zones_sub_dup_use %>%
  mutate(Dry.Tons.Use.State = use.allocator*Dry.Tons.Use.egrid_agg)

#write.csv(supplies_zones_sub_dup_use, "supplies_zones_sub_dup_use.csv")
```

## Get national sums by Resource
```{r}
national_resource <- supplies_zones_sub_dup_use %>%
  group_by(Resource, Resource.Form, Run.Name, Year) %>%
  summarise(Dry.Tons.ByResource = sum(Dry.Tons.Use.State))

write.csv(national_resource, "national_resource.csv")
```

